---
title: "Birth Country Imputation"
subtitle: "Advanced R for Econometricians"
type: "Term Project"
author: "Jamie Christy (256535) & Danaia Burtseva (237132)"
discipline: "MSc. Econometrics"
date: "`r Sys.Date()`"
studid: "256535 & 237132"
supervisor: "Prof. Dr. Martin Christopher Arnold"
secondsupervisor: "M.Sc. Martin Schmelzer"
# ssemester: 1
semester: "Summer Term 2025"
deadline: "September 8, 2025"
output:
  pdf_document:
    keep_tex: yes
    template: template.tex
    fig_caption: yes
    citation_package: biblatex
    number_sections: true
toc: true
lot: true
lof: true
graphics: true
tables: true
biblio-title: References
fontsize: 11pt
geometry: lmargin=4cm,rmargin=3.5cm,tmargin=2.5cm,bmargin=2.5cm
biblio-files: references.bib
classoption: a4paper
language: english
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gt)
library(dplyr)
library(broom)
```

# Introduction

Missing or non-standard data poses a persistent challenge in empirical research, but can sometimes be ameliorated. This paper demonstrates a practical implementation of imputing missing demographic information using auxiliary features and external reference datasets. 

We consider a set of mock administrative data in which 40% of records are missing country of birth, and implement a procedure to assign each of these a 'best guess' country. The challenges are broadly twofold: we must handle inconsistent formatting in birth city names, and we must match birth cities or citizenships to countries.

Our approach combines multiple imputation strategies in a hierarchical framework, beginning with the most reliable sources (exact city matches in official databases) and progressively incorporating less certain approaches (citizenship-based imputation) for remaining missing values. We use direct indications of country in birth city names, matches to German or international reference datasets, cities with some non-missing values, typical citizenships associated with cities, and a few manual mappings to assign cities to countries. When direct city matching fails, we fall back to record-level indication of citizenship, which generally but not always leads to identification of 'Germany.'

The majority of incomplete records are matched to a German city by direct reference, and many are matched using appearances of their birth city in the complete records, then we usually fall back to citizenship. Post-imputation, we see a population born 29% in Germany, 9% in Poland, 8% in Syria, and a long tail of other countries.

The remainder of this paper is structured as follows. Section \ref{sec:data} describes the data sources and the preprocessing to relate them to each other. Section \ref{sec:imputation} details our imputation methodology, including the hierarchical matching process and the stepwise impacts on the incomplete records. Section \ref{sec:results} presents the results, including final success rates and a visualization of the geographic distribution of imputed birth countries. Section \ref{sec:conclusion} discusses the limitations of our approach, potential extensions, and general takweaways for similar data imputation challenges.

# Data {#sec:data}

```{r data, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load and prepare summary statistics
raw_df <- read.csv("data/birthdata.csv", sep=";")
raw_ct <- nrow(raw_df)
mi_ct <- sum(is.na(raw_df$birth_country))
mi_pct <- round(mi_ct/raw_ct * 100, 1)
deu_ct <- sum(raw_df$citizenship_1 == "0" | raw_df$citizenship_2 == "0", na.rm = TRUE)
deu_pct <- round(deu_ct/raw_ct * 100, 1)

NNN<-0
```

Our work combines a given incomplete 'birth data' dataset with several external reference sources to enable systematic imputation of missing birth country information:
- country codes from Statistisches Bundesamt (Destatis)
- modern “Addresses of Municipal and City Administrations in Germany" from Die Statistischen Ämter des Bundes und der Länder
- historical eastern territories of Germany from NNNTODO
- major modern cities from the World Cities Database

## Primary dataset

Our dataset contains `r format(raw_ct, big.mark = ",")` observations of individuals with citizenship and birth location information, of which approximately `r format(mi_ct, big.mark = ",")` records (`r mi_pct`%) are missing birth country data. At a glance, the records show a strong German orientation, with `r deu_pct`% of all records marking a German citizenship and many incomplete records indicating German birth cities. In fact, there are no complete records with a German (000) birth country marked, suggesting a systematic null for German country of birth.

```{r raw_summary}
# raw_df %>% 
#   summarise(
#     `Total Records` = n(),
#     `Missing Birth Country` = sum(is.na(birth_country)),
#     `Missing Birth City` = sum(is.na(birth_city)),
#     `German Citizenship` = sum(citizenship_1 == "0" | citizenship_2 == "0", na.rm = TRUE)
#   ) %>%
#   knitr::kable(caption = "Dataset Overview")
```

Given the expectation of substantial German origin for the incomplete records, we focus on understanding and identifying German (or formerly German) cities, and prioritize European cities in cases of ambiguity, while maintaining global coverage for edge cases.

## External Reference Sources

We integrate four primary external datasets to enable comprehensive geographic matching:

### Destatis

The German Federal Statistical Office (*Statistisches Bundesamt*) provides standardized country and nationality codes through their "Staatsangehörigkeitsgebietsschlüssel" [@destatis2024]. This table (hereafter "DESTATIS codes") maps three-digit country codes to modern country names, nationalities, and ISO codes. We use this primarily to translate numeric country codes in our data to readable country names and to ensure consistent country assignments that align with German administrative standards.

### German Municipal Listing

The Statistical Offices of the Federation and States (*Die Statistischen Ämter des Bundes und der Länder*) publish an annual directory of municipal and city administrations [@statsaemter2023]. This "Municipal Directory" contains comprehensive listings of German cities and municipalities as of January 2023. We extract city names from both the official seat locations and administrative designations, providing a list of approximately `r NNN` potential names for German localities, including variations. This serves as our primary reference for identifying German birth cities, with both exact matching and simplified matching (removing common prefixes and administrative suffixes).

### Ostgebeit

For historical German territories now part of other countries, we manually compile a reference list of major cities in former German regions including East Prussia, Silesia, and Pomerania. This "Ostgebiete reference" handles cities like Königsberg (now Kaliningrad) or Breslau (now Wrocław) that may appear in historical records but no longer exist under German administration.

### International

We incorporate the World Cities Database from SimpleMaps [@simplemaps2024], which provides comprehensive global city coverage with population weights and country assignments. This "World Cities DB" contains over 40,000 cities worldwide and serves as our fallback reference for non-German locations. While its primary contribution is relatively minor (covering roughly `r NNN`% of our otherwise unmatched cases), it provides important validation that remaining unmatched records do not correspond to major international cities, supporting our assumption that most represent German locations with non-standard spellings.

```{r data-overview, results='asis'}
# Summary table of data sources and their contributions
overview_table <- data.frame(
  Metric = c("Total Records", "Missing Birth Country", "German Citizenship"),
  Count = c(
    format(raw_ct, big.mark = ","),
    format(mi_ct, big.mark = ","),
    format(deu_ct, big.mark = ",")
  ),
  Percentage = c(
    "100%",
    paste0(mi_pct, "%"),
    paste0(deu_pct, "%")
  )
)

# knitr::kable(overview_table, caption = "Dataset Overview")
```

The combination of these sources clarifies and prioritizes the expected German-heavy geographic distribution of our records, while also covering basic international and historical considerations.

# Imputation

There are a number of factors we can leverage to impute missing information:
- many records are complete, and can be used to identify cities or patterns.
- many incomplete records are likely to be German, and to follow naming patterns used in German administration.
- some records directly indicate a country or a well-known modern city.

Generally, we try to identify the birth country based on birth city. This is complicated by several factors that make birth city names non-trivial to attach to a country.
- birth city names often have non-standardized formats, inconsistent transliterations, or misspellings.
- many birth city names include a slash ("/") which *often but not always* indicates a sub/super unit of geography (typically City/District).
- birth cities may indicate a historical name, or a region rather than a city.

The imputation process follows a hierarchical progression of approaches to identification, beginning with the most reliable sources (explicitly named countries, exact city matches in official databases) and progressively incorporating less certain approaches (citizenship-based imputation) for remaining missing values. More precisely, we impute missing values in order, based on the following:


The detailed process is described in the rest of this section.

## Standardizing data

Necessary for grouping cities, and for matching to datasets.

We first standardize birth city names, splitting entries on delimiter characters, applying fuzzy matching techniques to handle minor spelling variations or errors, and parsing out direct indications of country. We then systematically match cities against reference datasets for German municipalities and international cities. 

## Named countries

## German cities

Here, we incorporate expected names and formatting of German administrative geography and historical political boundaries. For German cities, we implement both exact matching and simplified matching that removes common prefixes and suffixes. For international locations, we prioritize modern political boundaries for consistency. 

## Non-missing city-level information

For cities that appear in both complete and incomplete records, we extract empirical patterns from the complete data to inform imputation. We identify all city-country combinations in records with non-missing birth country information, calculating the frequency with which each city corresponds to each country. Cities where a single country accounts for more than 60% of complete observations are assigned that dominant country. For example, if "BELGRAD" appears in 10 complete records and 8 of those indicate Serbia as the birth country, we assign Serbia to any incomplete records listing "BELGRAD" as the birth city. This approach leverages the assumption that individuals sharing the same birth city are likely to share the same birth country, while accounting for cases where cities may legitimately correspond to multiple countries due to historical boundary changes or data entry variations. The 60% threshold balances the desire to use empirical patterns against the risk of misclassification in ambiguous cases.

```{r pressure, echo=FALSE, fig.cap="\\label{fig:pressure} Pressure", fig.height=4, fig.pos="t"}
plot(pressure)
```

Looking at Figure \ref{fig:pressure} makes me happy.

# Results

## The gt package

In empirical work, it's crucial not only to present your results but also to explain your research strategy. Often that involves presenting data and results in a tabular format. The `gt` package provides a modern and flexible way to generate high-quality LaTeX or HTML tables from R objects.

### Showing data using gt

```{r, message=FALSE, warning=FALSE, results='asis'}
data(trees)

n <- 6
trees_sample <- trees %>%
  slice_sample(n = n) %>%
  mutate(Row = row_number()) %>%
  relocate(Row)

gt_latex <- trees_sample %>%
  gt() %>%
  tab_header(
    title = md(
      paste(n, "Observations from the `trees` Dataset")
      )
  ) %>%
  cols_label(
    Girth = "Diameter",
    Height = "Height",
    Volume = "Volume"
  ) %>%
  tab_options(
    latex.tbl.pos = "H", 
    latex.use_longtable = FALSE
  ) %>%
  as_latex()

# Replace \begin{table} with \begin{table}[H]
gt_latex <- gsub(
  "\\\\begin\\{table\\}", 
  "\\\\begin{table}[H]", 
  gt_latex
)

cat(gt_latex)
```

### Regression summary with `gt`

```{r, message=FALSE, warning=FALSE, results='asis'}
mod <- lm(Volume ~ Height, data = trees)
mod_summary <- tidy(mod)

gt_latex <- mod_summary %>%
  mutate(term = recode(
    term, "(Intercept)" = "Intercept")
    ) %>%
  gt() %>%
  tab_header(
    title = "OLS Regression: Volume ~ Height"
  ) %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 3
  ) %>%
  cols_label(
    term = "Term",
    estimate = "Estimate",
    std.error = "Std. Error",
    statistic = "t Statistic",
    p.value = "p-Value"
  ) %>%
  as_latex()

# Replace \begin{table} with \begin{table}[H]
gt_latex <- gsub(
  "\\\\begin\\{table\\}", 
  "\\\\begin{table}[H]", 
  gsub(
    "caption\\*", 
    "caption", 
    gsub(
      "\\\\begin\\{tabular\\*\\}", 
      "\\\\label{tab:olsrestab}\n\\\\begin{tabular*}", 
      gt_latex
    )
  )
)

cat(gt_latex)
```

Table \ref{tab:olsrestab} is a table with a label and thus can be referenced. Captioned tables will appear in the LOT.

### Visualizing the model

```{r, fig.cap="Dataset and regression"}
plot(trees$Volume ~ trees$Height,
     xlab = "Height", 
     ylab = "Volume",
     main = "Linear Regression: Volume ~ Height")
abline(mod, col = "blue", lwd = 2)
```

# Conclusion

Missing data poses a persistent challenge in empirical research, particularly when key demographic variables are incomplete. This paper addresses the practical problem of imputing missing birth country information using auxiliary features and external reference datasets. In administrative and survey data, birth country information is frequently incomplete due to data entry errors, respondent non-response, or historical changes in data collection practices. The ability to accurately impute these missing values has important implications for demographic analysis, migration studies, and policy research.

This work contributes to the literature on practical data imputation in several ways. First, it demonstrates how external reference data can be systematically integrated into the imputation process, incorporating knowledge about geographic and political relationships. Second, it provides a framework for handling the common problem of non-standardized location names in administrative data. Third, it illustrates how auxiliary variables like citizenship can be used probabilistically when their relationship to the target variable is imperfect but informative.

A *bibtex* bibliography (see the fil `references.bib`) can be used for citations. The file name of the bibliography used in this template is `references.bib`.

You can cite a source like this: [@Keil_2012] or @Litterman1986.

A cited source will be automatically added to the end of the document.


\pagebreak

# Software References {-}
\addcontentsline{toc}{section}{Software References}

\newpage

# Appendix {-}
\addcontentsline{toc}{section}{Appendix}

## A.1 Description of relevant variables

## A.2 Fuzzy and manual matches
